---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
metadata <- read.csv("Data/airway_metadata.csv", stringsAsFactors = FALSE)
counts <- read.csv("Data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
```

```{r}
dim(counts)
head(counts)
```

```{r}
summary(counts)
```


```{r}
head(metadata)
```

```{r}
colnames(counts)
```

```{r}
metadata$id
```
Check for correspondence of metadata and countdata
```{r}
all(colnames(counts)[-1] == metadata$id)
```

First task find which columns in **counts** correspond to the control conditions (i.e. no drugs!). Then we calculate the mean in count value

```{r}
control <- metadata$dex == "control"
control.ids <- metadata$id[control.inds]
```

calculate the mean counts of all genes
```{r}
control.mean <- rowSums(counts[ ,control.ids])/ ncol(counts[,control.ids])
names(control.mean) <- counts$ensgene
head(control.mean)
```

Now do the same thing for the drug treated sample (i.e. **treated** columns)
```{r}
treated <- metadata[metadata[,"dex"] == "treated",]
treated.mean <- rowSums(counts[,treated$id]/ncol(counts[,treated$id]))
treated.ids <- metadata$id[control.inds]
names(treated.mean) <- counts$ensgene
```


```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

```{r}
colSums(meancounts)
```

```{r}
plot(meancounts$control.mean, meancounts$treated.mean)
```

Put this into a log log scale to visualize
```{r}
plot(log = "xy", meancounts$control.mean, meancounts$treated.mean)
```


```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts) 
```

```{r}
toy <- c(1,2,0,4,8)
toy == 0
```

find zero entry genes in meancounts
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)
to.rm <- unique(zero.vals[,1])
```


remove zero entry genes from meancounts
```{r}
mycounts <- meancounts[-to.rm,]
head(mycounts)
dim(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```


number of up and down regulated genes
```{r}
sum(up.ind)
sum(down.ind)
```

```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```

Use the merge function to add annotation data from ensemble
```{r}
mycounts.annotated <- merge(x = mycounts, y = anno, by.x = "row.names", by.y = "ensgene")
head(mycounts.annotated)
```

```{r}
library(DESeq2)
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds

```


```{r}
dds <- DESeq(dds)
```


```{r}
result1 <- results(dds, alpha = 0.01)
head(result1)
```


```{r}
summary(result1)
```

plot the result

```{r}
mycols <- rep("gray", nrow(result1))
mycols[abs(result1$log2FoldChange) >2]<-"blue" 
plot(result1$log2FoldChange, -log(result1$padj), col = mycols) 
```



```{r}

```


