---
title: "Class11: Structural Bioonformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Section 1. The PDB
```{r}
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats
```
Determine the percentage of structures solved by X-Ray and Electron Microscopy.
```{r}

```

From the website what proportion of structures are protein?
```{r}
percent.by.name <- stats$Total/sum(stats$Total)*100
names(percent.by.name) <- rownames(stats)
percent.by.name
sum(stats$Proteins)/sum(stats$Total)*100
```
## Section 3. Introduction to Bio3D in R
First load the package
```{r}
library(bio3d)
```

Now we can use the function from this bio3d package
```{r}
pdb <- read.pdb("1hsg")
pdb
```

```{r}
aa321(pdb$atom$resid)
```

select calpha
```{r}
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```

select protein
```{r}
prot.inds <- atom.select(pdb, "protein")
```

select ligand
```{r}
lig.inds <- atom.select(pdb,"ligand")
```

Use the Bio3D function to write out a protein only PDB file for viewing in VMD. 
```{r}
prot.pdb <- trim.pdb(pdb, prot.inds)
write.pdb(prot.pdb, file="protein.pdb")
```

Let's do the same thing for drug
```{r}
ligand.pdb <- trim.pdb(pdb, lig.inds)
write.pdb(ligand.pdb, file = "ligand.pdb")
```

