---
title: "Class09"
author: "Jiaqian Zhong"
date: "2/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
View(wisc.df)
```

How many samples (e.g patients) are in this data set?
```{r}
nrow(wisc.df)
```
How many variables/features in the data are suffixed with _mean?
```{r}
inds <- grep("_mean",colnames(wisc.data))
length(inds)
```

Convert the features of the data: wisc.data to a matrix
```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)
```

Set the row names of wisc.data to patient id
```{r}
row.names(wisc.data) <- wisc.df$id
#head(wisc.data)
```

Finally, setup a separate new vector called diagnosis to be 1 if a diagnosis is malignant ("M") and 0 otherwise. Note that R coerces TRUE to 1 and FALSE to 0
```{r}
table(wisc.df$diagnosis)
tmp <- rep(0, nrow(wisc.df))
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
cbind(diagnosis, wisc.df$diagnosis)
```

How many of the observations have a malignant diagnosis?
```{r}
sum(diagnosis)
```

Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled.
```{r}
colMeans(wisc.data)
apply(wisc.data,2,sd)
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)? 0.4427
Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data? PC1 + PC2 + PC3
Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data? PC1...PC7

```{r}
biplot(wisc.pr)
```

Now we need to make our own PCA plot
```{r}
# Scatter plot observations by components 1 and 2
plot(wisc.pr$x [,1], wisc.pr$x[,2], col = diagnosis+5,  xlab = "PC1", ylab = "PC2")
```

Calculate the variance of each principal component by squaring the sdev component of wisc.pr
```{r}
variance <- wisc.pr$sdev^2
pve <- round((variance/sum(variance))*100, 1)
```

```{r}
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained",  type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve), sep = ""), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

Hierarchical clustering
Scale the data
```{r}
data.scaled <- scale(wisc.data)
round(apply(wisc.data, 2, sd), 1)
```

Calculate the (Euclidean) distances between all pairs of observations in the new scaled dataset and assign the result to data.dist.
```{r}
data.dist <- hist(data.scaled)
```

Create a hierarchical clustering model using complete linkage. 
```{r}
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "complete")
```

```{r}
plot(wisc.hclust)
```

Cluster in PCA result
For clustering, we need: 
1. distance matrix
2. Clustering method
3. Cutree
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+5)
```

```{r}
pc.dist <- dist(wisc.pr$x[,1:2])
pc.hclust <- hclust(pc.dist, method = "ward.D2")
plot(pc.hclust)
```

```{r}
grsp3 <- cutree(pc.hclust, k=2)
table(grsp3)
```

```{r}
table(grsp3, diagnosis)
```

We will use the predict() function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=grsp3)
points(npc[,1], npc[,2], col="blue", pch=16)
```

